# Time series decomposition

```{r, include = FALSE}
# Libraries
library(tidyverse)
library(tsibble)
library(fpp3)
library(knitr)
library(svglite)
library(ggthemes)
library(lemon)
library(USgas)

# Datasets
olympic_running <- tsibbledata::olympic_running
pbs <- tsibbledata::PBS
vic_elec <- tsibbledata::vic_elec
tourism <- tsibble::tourism
aus_production <- tsibbledata::aus_production
gafa_stock <- tsibbledata::gafa_stock
pelt <- tsibbledata::pelt
aus_arrivals <- fpp3::aus_arrivals
aus_retail <- tsibbledata::aus_retail
us_employment <- fpp3::us_employment
global_economy <- tsibbledata::global_economy
aus_livestock <- tsibbledata::aus_livestock

# Functions

## Get the lambda for a Box-Cox transformation
## using the Guerrero method
get_boxcox_lambda <- function(df, var) {
  df %>%
  features(df[, var], features = guerrero) %>%
  pull(lambda_guerrero)
}

# Parameters
kable_length <- 5
svg_res <- 144

# GGplot graphical style
theme_set(
  theme_tufte() +
  theme(
    axis.line = element_line(colour = "black", size = rel(1))
  )
)
short_axis <- function(baxis = "both", laxis = "both") {
  return(lemon::coord_capped_cart(bottom = baxis, left = laxis))
}
```

## Transformations and adjustments

```{r}
global_economy %>%
  filter(Country == "Australia") %>%
  autoplot(GDP / Population) +
  labs(title = "GDP per capita", y = "$US")
```

```{r}
print_retail <- aus_retail %>%
  filter(Industry == "Newspaper and book retailing") %>%
  group_by(Industry) %>%
  index_by(Year = year(Month)) %>%
  summarise(Turnover = sum(Turnover))
aus_economy <- global_economy %>%
  filter(Code == "AUS")
```

```{r}
print_retail %>%
  left_join(aus_economy, by = "Year") %>%
  mutate(Adjusted_turnover = Turnover / CPI * 100) %>%
  pivot_longer(c(Turnover, Adjusted_turnover),
               values_to = "Turnover") %>%
  mutate(name = factor(name,
         levels = c("Turnover", "Adjusted_turnover"))) %>%
  ggplot(aes(x = Year, y = Turnover)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y") +
  labs(title = "Turnover: Australian print media industry",
       y = "$AU")
```

```{r}
lambda <- aus_production %>%
  features(Gas, features = guerrero) %>%
  pull(lambda_guerrero)
aus_production %>%
  autoplot(box_cox(Gas, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed gas production with $\\lambda$ = ",
         round(lambda, 2))))
```

## Time series components

```{r}
us_employment
```

```{r}
us_retail_employment <- fpp3::us_employment %>%
  filter(year(Month) >= 1990, Title == "Retail Trade") %>%
  select(-Series_ID)
autoplot(us_retail_employment, Employed) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
```

```{r}
dcmp <- us_retail_employment %>%
  model(stl = STL(Employed))
components(dcmp)
```

```{r}
components(dcmp) %>%
  as_tsibble() %>%
  autoplot(Employed, colour = "gray") +
  geom_line(aes(y = trend), colour = "#D55E00") +
  labs(
    y = "Persons (thousands)",
    title = "Total employment in US retail"
  )
```

```{r}
components(dcmp) %>% autoplot()
```

```{r}
components(dcmp) %>%
  as_tsibble() %>%
  autoplot(Employed, colour = "gray") +
  geom_line(aes(y = season_adjust), colour = "#0072B2") +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
```

## Moving averages

```{r}
global_economy
```

```{r}
global_economy %>%
  filter(Country == "Australia") %>%
  autoplot(Exports) +
  labs(y = "% of GDP", title = "Total Australian exports")
```

```{r}
aus_exports <- global_economy %>%
  filter(Country == "Australia") %>%
  mutate(
    `5-MA` = slider::slide_dbl(Exports, mean,
                .before = 2, .after = 2, .complete = TRUE)
  )
```

```{r}
aus_exports %>%
  autoplot(Exports) +
  geom_line(aes(y = `5-MA`), colour = "#D55E00") +
  labs(y = "% of GDP",
       title = "Total Australian exports") +
  guides(colour = guide_legend(title = "series"))
```

```{r}
beer <- aus_production %>%
  filter(year(Quarter) >= 1992) %>%
  select(Quarter, Beer)
beer_ma <- beer %>%
  mutate(
    `4-MA` = slider::slide_dbl(Beer, mean,
                .before = 1, .after = 2, .complete = TRUE),
    `2x4-MA` = slider::slide_dbl(`4-MA`, mean,
                .before = 1, .after = 0, .complete = TRUE)
  )
```

```{r}
beer_ma %>%
  autoplot(Beer, colour = "gray") +
  geom_line(aes(y = `4-MA`), colour = "#D55E00") +
  geom_line(aes(y = `2x4-MA`), colour = "#0039d5")
```

```{r}
us_retail_employment_ma <- us_retail_employment %>%
  mutate(
    `12-MA` = slider::slide_dbl(Employed, mean,
                .before = 5, .after = 6, .complete = TRUE),
    `2x12-MA` = slider::slide_dbl(`12-MA`, mean,
                .before = 1, .after = 0, .complete = TRUE)
  )
us_retail_employment_ma %>%
  autoplot(Employed, colour = "gray") +
  geom_line(aes(y = `2x12-MA`), colour = "#D55E00") +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail")
```

## Classical decomposition

```{r}
us_retail_employment %>%
  model(
    classical_decomposition(Employed, type = "additive")
  ) %>%
  components() %>%
  autoplot() +
  labs(title = "Classical additive decomposition of total
                  US retail employment")
```

## Better methods

```{r}
x11_dcmp <- us_retail_employment %>%
  model(x11 = X_13ARIMA_SEATS(Employed ~ x11())) %>%
  components()
autoplot(x11_dcmp) +
  labs(title =
    "Decomposition of total US retail employment using X-11.")
```

```{r}
x11_dcmp %>%
  ggplot(aes(x = Month)) +
  geom_line(aes(y = Employed, colour = "Data")) +
  geom_line(aes(y = season_adjust,
                colour = "Seasonally Adjusted")) +
  geom_line(aes(y = trend, colour = "Trend")) +
  labs(y = "Persons (thousands)",
       title = "Total employment in US retail") +
  scale_colour_manual(
    values = c("gray", "#0072B2", "#D55E00"),
    breaks = c("Data", "Seasonally Adjusted", "Trend")
  )
```

```{r}
x11_dcmp %>%
  gg_subseries(seasonal)
```

```{r}
seats_dcmp <- us_retail_employment %>%
  model(seats = X_13ARIMA_SEATS(Employed ~ seats())) %>%
  components()
autoplot(seats_dcmp) +
  labs(title =
    "Decomposition of total US retail employment using SEATS")
```

## STL decomposition

```{r}
us_retail_employment %>%
  model(
    STL(Employed ~ trend(window = 7) +
                   season(window = "periodic"),
    robust = TRUE)) %>%
  components() %>%
  autoplot()
```

`window = 'periodic` restricts seasonality to be identical over the years. 
A shorter window allows for faster changes in the trend and in the seasonality.
This is necessary here, notably to see

## Exercises

1. Countries GDP

```{r}
excluded_groups <- c("World", "Arab World",
  "Central Europe and the Baltics", "Early-demographic dividend",
  "East Asia & Pacific", "East Asia & Pacific (excluding high income)",
  "East Asia & Pacific (IDA & IBRD countries)", "Euro area",
  "Europe & Central Asia", "Europe & Central Asia (excluding high income)",
  "Europe & Central Asia (IDA & IBRD countries)", "European Union",
  "Fragile and conflict affected situations",
  "Heavily indebted poor countries (HIPC)", "High income",
  "IBRD only", "IDA & IBRD total", "IDA blend", "IDA only",
  "IDA total", "Late-demographic dividend", "Latin America & Caribbean",
  "Latin America & Caribbean (excluding high income)",
  "Latin America & the Caribbean (IDA & IBRD countries)",
  "Least developed countries: UN classification",
  "Low & middle income", "Low income", "Lower middle income",
  "Middle East & North Africa", "Middle East & North Africa (excluding high income)",
  "Middle East & North Africa (IDA & IBRD countries)", "Middle income",
  "North America", "OECD members", "Other small states", "Pacific island small states",
  "Post-demographic dividend", "Pre-demographic dividend", "Small states",
  "South Africa", "South Asia", "South Asia (IDA & IBRD)",
  "Sub-Saharan Africa", "Sub-Saharan Africa (excluding high income)",
  "Upper middle income"
  )
```

```{r}
global_economy %>%
  filter(!Country %in% excluded_groups) %>%
  ggplot(
    aes(x = Year, y = GDP / Population, color = Country)
  ) +
  geom_line() +
  theme(legend.position = "none")
```


```{r}
as_tibble(global_economy) %>%
  drop_na(GDP) %>%
  filter(!Country %in% excluded_groups) %>%
  mutate(gdp_capita = GDP / Population) %>%
  group_by(Year) %>%
  filter(gdp_capita == max(gdp_capita)) %>%
  select(Country, Year, gdp_capita) %>%
  arrange(Year) %>%
  rename(`GDP per capita` = gdp_capita) %>%
  kable()
```

Highest GDP per capita: mostly very small (and rich) countries.

2. Transformations

Note: I'd be a bit [careful with transformations](https://stats.stackexchange.com/questions/18844/when-and-why-should-you-take-the-log-of-a-distribution-of-numbers).
Maybe do this based on an analysis of residuals from a regression, in the future. Plus, these transformations won't be as interpretable as logs.

Remember that the transformation is done to make the magnitude of the seasonal variation more uniform.

```{r}
global_economy %>%
  filter(Country == "United States") %>%
  autoplot(GDP)
```

```{r}
global_economy %>%
  filter(Country == "United States") %>%
  mutate(
    GDP = box_cox(
      x = GDP,
      lambda = get_boxcox_lambda(., "GDP")
    )
  ) %>%
  autoplot(GDP)
```

We get a slightly more "linear" variable. I'd rather just take the log.

```{r}
aus_livestock %>%
  filter(Animal == "Bulls, bullocks and steers", State == "Victoria") %>%
  autoplot(Count)
```

Not sure a transformation will be useful here, but I may be wrong.

```{r}
aus_livestock %>%
  filter(Animal == "Bulls, bullocks and steers", State == "Victoria") %>%
  mutate(Count = box_cox(
    x = Count,
    lambda = get_boxcox_lambda(., "Count")
    )
  ) %>%
  autoplot(Count)
```

Probably useless here.

```{r}
vic_elec %>%
  autoplot(Demand)
```

No point in doing anything here. There does not seem to be any trend, and the variable is highly non-linear.

```{r}
vic_elec_stl <- vic_elec %>%
  model(stl = STL(Demand)) %>%
  components(stl)

vic_elec_stl %>%
  autoplot()
```

Quite complex.

```{r}
aus_production %>%
  autoplot(Gas)
```

```{r}
aus_production %>%
  mutate(
    Gas = box_cox(
      x = Gas,
      lambda = get_boxcox_lambda(., "Gas")
    )
  ) %>%
  autoplot(Gas)
```

Not sure why the transformation does not help here.

4. Retail

```{r}
set.seed(451)
aus_retail_sample <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`, 1))
```
```{r}
aus_retail_sample %>%
  autoplot(Turnover)
```

Box-cox should be useful there.

```{r}
aus_retail_sample %>% get_boxcox_lambda(., "Turnover")

aus_retail_sample %>%
  mutate(Turnover = log(Turnover)
  ) %>% 
  autoplot(Turnover)
```

Lambda is nearly identical to 0. I took the log instead.

5. I'll skip the Box-Cox transformations. I've done enough.

6. 3x5-MA = 7-MA with weights

5-MA:

$$
MA_t = \frac{1}{5}(y_{t-2} + y_{t-1} + y_t + y_{t+1} + y_{t+2})
$$

3x5-MA:

$$
\frac{1}{3}(\frac{1}{5}MA_{t-1} + \frac{1}{5}MA_{t} + \frac{1}{5}MA_{t+1}) \\
= \frac{1}{15}MA_{t-1} + \frac{1}{15}MA_{t} + \frac{1}{15}MA_{t+1} \\
= \frac{1}{15}y_{t-3} + \frac{2}{15}y_{t-2} + \frac{3}{15}y_{t-1} + \frac{3}{15}y_t + \frac{3}{15}y_{t+1} + \frac{2}{15}y_{t+2} + \frac{1}{15}y_{t+3}
$$

7. Some plots on aus_production

```{r}
gas <- tail(aus_production, 5*4) %>% dplyr::select(Gas)
```

```{r}
gas %>% autoplot(Gas)
```

Increasing trend. Low output (or demand) in Q1 and Q4, high in Q2 & Q3.

```{r}
gas %>%
  model(
    classical_decomposition(Gas, type = "multiplicative")
  ) %>%
  components %>%
  autoplot()
```

I was right.

```{r}
gas %>%
  model(
    classical_decomposition(Gas, type = "multiplicative")
  ) %>%
  components %>%
  select(Quarter, Gas, season_adjust) %>%
  autoplot(Gas, colour = "gray") +
  geom_line(aes(y=season_adjust), colour = "red")
```

```{r}
gas %>%
  mutate(
    Quarter_char = Quarter %>% as.character(),
    Gas = replace(
      Gas,
      Quarter_char == "2008 Q1",
      Gas[Quarter_char == "2008 Q1"] + 300
    )
  ) %>%
  model(
    classical_decomposition(Gas, type = "multiplicative")
  ) %>%
  components %>%
  autoplot()
```

An outlier does mess things up. Since the outlier is in Q1, now the seasonal demand in Q1 is very high. Plus, the outlier is seen in the trend, and in the random component.

```{r}
gas %>%
  mutate(
    Quarter_char = Quarter %>% as.character(),
    Gas = replace(
      Gas,
      Quarter_char == "2006 Q1",
      Gas[Quarter_char == "2006 Q1"] + 300
    )
  ) %>%
  model(
    classical_decomposition(Gas, type = "multiplicative")
  ) %>%
  components %>%
  autoplot()
```

Still problematic. But it doesn't seem to change much, compared to an outlier in the middle in of time series.

8. X-11 on retail data

```{r}
aus_retail_sample %>%
  autoplot()
```

No transformation

```{r}
aus_retail_sample %>%
  model(
    x11 = X_13ARIMA_SEATS(Turnover ~ x11())
  ) %>%
  components %>%
  autoplot()
```

Log transformed

```{r}
aus_retail_sample %>%
  model(
    x11 = X_13ARIMA_SEATS(log(Turnover) ~ x11())
  ) %>%
  components %>%
  autoplot()
```

There doesn't seem to be anything weird. Except some irregularity in the remainder plot, on the beginning of the 2000s. But nothing out of the ordinary.

9. Skipped. Just describing graphs.

10. Canadian gas

```{r}
canadian_gas %>%
  autoplot()
```

```{r}
canadian_gas %>%
  gg_season()
```

Indeed: low seasonality in the 60s, quite higher in the mid-70S to 80s. There is still a bit of seasonality after that.

```{r}
canadian_gas %>%
  gg_subseries()
```

```{r}
canadian_gas %>%
  model(
    STL(Volume ~ trend(window = 18) +
                   season(window = 6),
    robust = TRUE)) %>%
  components() %>%
  autoplot()
```

Maybe there is a metric or some indicator to correctly choose the time window. A 6-month window allows us to see the evolution of seasonality over the years quite nicely.

```{r}
canadian_gas %>%
  model(
    STL(Volume ~ trend(window = 18) +
                   season(window = 6),
    robust = TRUE)) %>%
  components() %>%
  select(Month, season_year) %>%
  gg_season()
```

Nice. Production gets increasingly lower in February and November through the years. It seems to have increased in July and August (during colder months of winter, then). Sometimes, the variation is not linear: production was lower in the early 1960s, increased rapidly in the 70s to 80s, and then decreased a little bit (e.g. January and December)

```{r}
canadian_gas %>%
  model(
    STL(Volume ~ trend(window = 18) +
                   season(window = 6),
    robust = TRUE)) %>%
  components() %>%
  select(Month, season_adjust) %>%
  autoplot()
```

It seems ok. I have tried to change the window settings, but they don't seem to change the trend that drastically.

```{r}
canadian_gas %>%
  model(
    x11 = X_13ARIMA_SEATS(Volume ~ x11())
  ) %>%
  components() %>%
  autoplot()
```

Surprising. Here, seasonality is a lot higher in the sixties. 

```{r}
canadian_gas %>%
  model(
    x11 = X_13ARIMA_SEATS(Volume ~ seats())
  ) %>%
  components() %>%
  autoplot()
```

x11 and SEATS are quite similar to my untrained eyes, to be fair. I guess there are some subtleties in each that make them more efficient in specific situations. For now, both seem good.
See [Seasonal Adjustment Methods and Real Time Trend-cycle Estimation](https://www.amazon.fr/dp/3319318209?tag=otexts0e-21&geniuslink=true) for a discussion about it.

```{r}
canadian_gas %>%
  model(
    x11 = X_13ARIMA_SEATS(Volume ~ x11())
  ) %>%
  components() %>%
  select(Month, seasonal) %>%
  gg_season()
```

It's funny to see the seasonality is quite different from the one obtained with STL.