# Time series graphics

## Tsibble

```{r, include = FALSE}
# Libraries
library(tidyverse)
library(tsibble)
library(fpp3)
library(knitr)
library(svglite)
library(ggthemes)
library(lemon)
library(USgas)

# Datasets
olympic_running <- tsibbledata::olympic_running
PBS <- tsibbledata::PBS # nolint
vic_elec <- tsibbledata::vic_elec
tourism <- tsibble::tourism
aus_production <- tsibbledata::aus_production
gafa_stock <- tsibbledata::gafa_stock
pelt <- tsibbledata::pelt
aus_arrivals <- fpp3::aus_arrivals
aus_retail <- tsibbledata::aus_retail

# Parameters
kable_length <- 5
svg_res <- 144

# GGplot graphical style
theme_set(
  theme_tufte() +
  theme(
    axis.line = element_line(colour = "black", size = rel(1))
  )
)
short_axis <- function(baxis = "both", laxis = "both") {
  return(lemon::coord_capped_cart(bottom = baxis, left = laxis))
}
```

Tsibble example

```{r}
y <- tsibble(
  Year = 2015:2019,
  Observation = c(123, 39, 78, 52, 110),
  index = Year
)
y
```

Olympic running dataframe

```{r}
kable(olympic_running[1:kable_length, ])
olympic_running %>% distinct(Sex, Length) %>% kable()
```

Some training on dplyr

```{r}
kable(PBS[1:kable_length, ])
PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  slice_head(n = 5) %>%
  kable()

# Total cost by monthyear
PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalCost = sum(Cost))

# Total cost by monthyear (M dollars)
PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalCost = sum(Cost)) %>%
  mutate(TotalCost = TotalCost / 1e6) -> a10
a10
```

Importing a csv

```{r}
prison <- readr::read_csv("https://OTexts.com/fpp3/extrafiles/prison_population.csv") # nolint
prison

prison <- prison %>%
  mutate(Quarter = yearquarter(Date)) %>%
  select(-Date) %>%
  as_tsibble(key = c(State, Gender, Legal, Indigenous),
             index = Quarter)

prison
```

## Time plots

Ansett dataframe

```{r}
ansett <- ansett
ansett
```

```{r}
ansett %>%
  filter(Airports == "MEL-SYD", Class == "Economy") %>%
  mutate(Passengers = Passengers / 1000) -> melsyd_economy
melsyd_economy
```

```{r}
# For svg: https://www.r-bloggers.com/2020/05/superior-svg-graphics-rendering-in-r-and-why-it-matters/ # nolint
autoplot(melsyd_economy, Passengers) +
  labs(
    title = "Ansett airlines economy class",
    subtitle = "Melbourne-Sydney",
    y = "Passengers ('000)"
  ) + short_axis()
```


## Time series patterns

## Seasonal plots

```{r}
autoplot(a10, TotalCost) +
  labs(
    y = "$ (millions)",
    title = "Australian antidiabetic drug sales"
  )
```
```{r}
a10 %>%
  gg_season(TotalCost, labels = "both") +
  labs(y = "$ (millions)",
       title = "Seasonal plot: Antidiabetic drug sales")
```

### Electricity demand graphs

```{r}
as_tibble(vic_elec) %>%
  group_by(Date = floor_date(Time, "day")) %>%
  summarise(
    Demand = mean(Demand),
    Temperature = mean(Temperature)
  ) %>%
  as_tsibble(index = Date) %>%
  autoplot(Demand)
```

There is a peak at +7000MW at the beginning of 2014. Curious.

```{r}
as_tibble(vic_elec) %>%
  group_by(Date = floor_date(Time, "day")) %>%
  summarise(
    Demand = mean(Demand),
    Temperature = mean(Temperature)
  ) %>%
  filter(Demand > 6500)
```

14 & 15 Jan 2014. There was a [heatwave](https://www2.health.vic.gov.au/Api/downloadmedia/%7BDC381402-DF8F-42A5-8153-2BDF690F5402%7D).

```{r}
as_tibble(vic_elec) %>%
  group_by(Date = floor_date(Time, "month")) %>%
  summarise(Demand = mean(Demand)) %>%
  as_tsibble(index = Date) %>%
  autoplot(Demand)
```

Seasonal plots

```{r}
vic_elec %>% gg_season(Demand, period = "day") +
  theme(legend.position = "none") +
  labs(y = "MW", title = "Electricity demand: Victoria")
```

```{r}
vic_elec %>% gg_season(Demand, period = "year") +
  labs(y = "MW", title = "Electricity demand: Victoria")
```

## Seasonal subseries plots

```{r}
a10 %>%
  gg_subseries(TotalCost) +
  labs(
    y = "$ (millions)",
    title = "Australian antidiabetic drug sales"
  )
```

```{r}
holidays <- tourism %>%
  filter(Purpose == "Holiday") %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
holidays
```

```{r}
autoplot(holidays, Trips) +
  labs(y = "Overnight trips ('000)",
       title = "Australian domestic holidays")
```

```{r}
gg_season(holidays, Trips) +
  labs(y = "Overnight trips ('000)",
       title = "Australian domestic holidays")
```

```{r}
holidays %>%
  gg_subseries(Trips) +
  labs(y = "Overnight trips ('000)",
       title = "Australian domestic holidays")
```

## Scatterplots

```{r}
vic_elec %>%
  filter(year(Time) == 2014) %>%
  autoplot(Demand) +
  labs(y = "MW",
       title = "Half-hourly electricity demand: Victoria")
```

```{r}
vic_elec %>%
  filter(year(Time) == 2014) %>%
  autoplot(Temperature) +
  labs(
    y = "Degrees Celsius",
    title = "Half-hourly temperatures: Melbourne, Australia"
  )
```

```{r}
vic_elec %>%
  filter(year(Time) == 2014) %>%
  ggplot(aes(x = Temperature, y = Demand)) +
  geom_point() +
  labs(x = "Temperature (degrees Celsius)",
       y = "Electricity demand (MW)")
```

```{r}
visitors <- tourism %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
visitors %>%
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State), scales = "free_y") +
  labs(title = "Australian domestic tourism",
       y = "Overnight trips ('000)")
```

```{r}
visitors %>%
  pivot_wider(values_from = Trips, names_from = State) %>%
  GGally::ggpairs(columns = 2:9)
```

## Lag plots

```{r}
aus_production
```

```{r}
recent_production <- aus_production %>%
  filter(year(Quarter) >= 2000)
recent_production %>%
  gg_lag(Beer, geom = "point") +
  labs(x = "lag(Beer, k)")
```

## Autocorrelation
```{r}
recent_production %>% ACF(Beer, lag_max = 9)
recent_production %>%
  ACF(Beer) %>%
  autoplot() + labs(title = "Australian beer production")
```

```{r}
a10 %>%
  ACF(TotalCost, lag_max = 48) %>%
  autoplot() +
  labs(title = "Australian antidiabetic drug sales")
```

## White noise

```{r}
y <- tsibble(sample = 1:1000, wn = rnorm(1000), index = sample)
y %>% autoplot(wn) + labs(title = "White noise", y = "")
```

```{r}
y %>%
  ACF(wn) %>%
  autoplot() + labs(title = "White noise")
```

## Exercises

1. Use the help function to explore what the series gafa_stock, PBS, vic_elec and pelt represent.
  a. use autoplot() to find what days corresponded to the peak closing price for each of the four stocks in gafa_stock
  b. What is the time interval of each series

`gafa_stock`

Daily data. Except week-ends.
```{r}
gafa_stock
gafa_stock %>%
  autoplot(Open)
```
```{r}
gafa_stock %>%
  autoplot(Volume)
```
```{r}
gafa_stock %>%
  ggplot(
    aes(x = Open, y = Volume, color = Symbol)
  ) +
  geom_point() +
  labs(
    x = "Opening price",
    y = "Volume"
  ) +
  short_axis()
```
```{r}
gafa_stock %>%
  mutate(ValueTraded = Open * Volume / 1e9) %>%
  autoplot(ValueTraded)
```

Efficient solution to get maximum Close for each Symbol.

```{r}
as_tibble(gafa_stock) %>%
  group_by(Symbol) %>%
  filter(Close == max(Close)) %>%
  select(Symbol, Date, Close) %>%
  kable()
```

Inefficient solution 

```{r}
for (company in factor(gafa_stock$Symbol) %>% levels()) {
  gafa_stock %>%
    filter(Symbol == company) %>%
    arrange(desc(Close)) %>%
    print()
}
```

3. Download the file tute1.csv from the book website, open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

```{r}
tute1 <- readr::read_csv("tute1.csv") %>%
  mutate(Quarter = yearmonth(Quarter)) %>%
  as_tsibble(index = Quarter)
tute1
```

```{r}
tute1 %>%
  pivot_longer(!Quarter) %>%
  ggplot(
    aes(x = Quarter, y = value, color = name)
  ) +
  geom_line()
```

```{r}
tute1 %>%
  pivot_longer(!Quarter) %>%
  ggplot(
    aes(x = Quarter, y = value, color = name)
  ) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y")
```

4. `USgas`

```{r}
us_total <- USgas::us_total %>%
  as_tsibble(index = year, key = state)
```

```{r}
new_england <- c(
  "Maine", "Vermont", "New Hampshire", "Massachusetts",
  "Connecticut", "Rhode Island"
)

us_total %>%
  filter(state %in% new_england) %>%
  ggplot(
    aes(x = year, y = y, color = state)
  ) +
  geom_line() +
  labs(x = "Year", y = "Gas consumption")
```

5. Tourism.xlsx

```{r}
tourism_xl <- readxl::read_excel("tourism.xlsx") %>%
  mutate(Quarter = yearquarter(Quarter)) %>%
  as_tsibble(index = Quarter, key = c("Region", "State", "Purpose"))
```

```{r}
tourism_xl %>%
  group_by(Region, Purpose) %>%
  summarise(Trips = mean(Trips)) %>%
  arrange(desc(Trips))
```

```{r}
tourism_xl %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
```

6. Trivial. I have done that too much already.

7. `aus_arrivals` 

```{r}
aus_arrivals
```

```{r}
aus_arrivals %>%
  autoplot(Arrivals) +
  short_axis()
```

The magnitude of the seasonality is a lot higher for NZ and the UK. Less so for Japan and the US.
All countries seem to follow an increasing trend, except for flights coming from Japan: it peaked in the mid 90s, and has decreased steadily ever since.

```{r}
aus_arrivals %>%
  gg_season(y = Arrivals)
```

Arrivals from the UK (and the US?) are noticeably higher in Q1 and Q4. Understandable, since it's winter in the northern hemisphere.
For NZ, we observe the opposite.

```{r}
aus_arrivals %>%
  gg_subseries(y = Arrivals)
```

It seems arrivals from the UK actually began decreasing a few years ago.  
There are some peaks in arrivals from the US at the end of the 80s and the 90s.

8. `aus_retail`

```{r}
set.seed(451)
aus_retail_sample <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`, 1))
```
```{r}
aus_retail_sample %>%
  autoplot(Turnover)
```

Increasing turnover, alright. There is seasonality and an inrceasing trend, but there does not seem to be significant cyclicity.

```{r}
aus_retail_sample %>%
  gg_season(Turnover)
```

It gets noticeably higher in November & December. Also in March, and sometimes during the summer?

```{r}
aus_retail_sample %>%
  gg_subseries(Turnover)
```

Same observation.

```{r}
aus_retail_sample %>%
  gg_lag(Turnover)
```

Same observation as before : December has the highest turnover, compared to every other month.

```{r}
aus_retail_sample %>%
  ACF(Turnover) %>%
  autoplot()
```

Unsurprisingly, there is (very) significant positive autocorrelation.

12. Google

```{r}
dgoog <- gafa_stock %>%
  filter(Symbol == "GOOG", year(Date) >= 2018) %>%
  mutate(trading_day = row_number()) %>%
  update_tsibble(index = trading_day, regular = TRUE) %>%
  mutate(diff = difference(Close))
```

```{r}
dgoog %>%
  ACF(Close) %>%
  autoplot()
```

```{r}
dgoog %>%
  ACF(diff) %>%
  autoplot()
```

It does seem to be white noise.